\prefacesection{Acknowledgments}

Getting to know Prof. Dan Jurafsky is the most precious part of my experience at Stanford. He perfectly exemplifies what a supportive and compassionate person should be like. 
Over the years, I have always been taking what he did for me
for granted: taking for granted that he spent 5 hours on 
editing
every paper and correcting all the typos and grammar mistakes from a non-native English speaker; taking for granted his generous support 
each and 
every time when I experienced down times in graduate school; taking for granted 
that he is always flexible with my schedules such as going back to 
the
mainland China visiting families in the middle of the school year. The opportunity of working with him and getting to know him is the most valuable part of my years at Stanford, and these experiences 
transcend all the tiny academic attainments in graduate school.

I want to thank 
Prof. Eduard Hovy. It is hard to evaluate how much his support means to me during the hardest time of my life. He was not like an advisor, but like an elder friend, 
telling an ignorant twenty-something how he should face ups and downs of life. He instilled in me the great importance of doing big work and solving big problems. Unfortunately, my pragmatic nature made me unwilling to invest time into ambitious, but risky research projects. I  kept working on tiny incremental projects to satisfy my utilitarian desire for publishing. Not living up to his expectation is the biggest regret of my graduate school experience. 


I want to particularly thank Prof. Claire Cardie. When I was a first-year graduate student studying biology at Cornell, barely having any experience with coding but trying to switch to the field of computer science,
she gave me the opportunity of being exposed to the intriguing nature of language processing using AI,  
and this opportunity
 completely changed the trajectory of my life. 
She invested her precious time into teaching me how to do research, and was patient as I slowly developed basic  coding skills. Her support continues over years after I left Cornell. 

I want to  thank Prof. Alan Ritter, for introducing me to 
 the basic skills in doing research.
I want to thank Prof. Sujian Li, for introducing me to the
beautiful 
 world of computer science, and for 
the support over the past many years.
I owe a big thank-you to Prof. Peter Doerschuk from the Biomedical Engineering department at Cornell, for all the encouragement and  support when I was trying to change my research field, and for telling me nothing is more important that doing things that I am happy about.  
 
During my years in graduate school, I have been blessed with the friendship of many amazing people. 
I want to thank all friends from Stanford IVgrad,  Daniel Heywood, Andrew Melton, 
Shannon Kao,
Ben Bell, Vivian Chen, Justin Su, Emily Kolenbrander, Tim Althoff, 
Caroline Abadeer,
Connor Kite
and Wendy Quay,
who provide me with a sense of family and identity in graduate school. 
IVgrad teaches me life is not all about research and research is not all about myself. 
Over the years, 
I have had the opportunity  of befriending, working with or learning from many wonderful students, professors
and researchers  at Stanford, CMU, Cornell, Microsoft Research, and Facebook, including 
Gabor Angeli, 
Chris Brockett, 
Danqi Chen, Xinlei Chen, Sumit Chopra, 
Eyal Cidon, 
 Bill Dolan, 
 Chris Dyer, 
 Jianfeng Gao, Michel Galley, 
Kelvin Guu,
He He, De-an Huang, Ting-Hao Kenneth Huang,
Sujay Kumar Jauhar,
Sebastien Jean, 
Percy Liang, 
Minh-Thang Luong, Hector Zhengzhong Liu, Max Xuezhe Ma, Chris Manning, 
Andrew Maas, 
Bill MacCartney,
Alexander Miller, 
Myle Ott, 
Chris Potts, 
Marc'aurelio Ranzato,
Mrinmaya Sachan,
Tianlin Shi, 
Noah Smith,
Russell Stewart, 
Yi-Chiang Wang, 
Di Wang, 
Willian Wang, 
Lu Wang, 
Sam Wiseman, 
Jason Weston, 
Ziang Xie,
Bishan Yang,
Diyi Yang, 
Shou-Yi Yu,
Arianna Yuan,
Jake Zhao and many many others who I am grateful for but cannot list all their names. 
I want to specially thank Will Monroe, Sida Wang, 
and Ziang Xie for the friendship, and
 for always being helpful and encouraging when
 graduate school 
  life is hard. 

I want to thank Prof. Emma Brunskill,
Prof. Christopher Potts  and Jason Weston, for agreeing to be on my thesis committee and  helping me all the way through the proposal, the defense and the 
final version of the thesis. I also want to thank Prof. Olivier Gevaert for helping chair my thesis defense. 

A special thank is owed to the administrative staff: Jay Subramanian, Helen Buendicho, Stacey Young, 
and Becky Stewart, who have help reduced many of the stresses in graduate school. 

Finally, I want to thank my families. My debt to them is unbounded.

